---
title: "Bank Marketing Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(highcharter)
library(gt)
library(htmltools)
library(viridisLite)
library(shiny)
library(shinydashboard)
```


```{r}
#importing data
df <- read_csv("Bank_Personal_loan_Modelling.csv")
df2 <- df[df$Experience >= 0,]

colnames(df2) <- gsub(" ", "_", colnames(df2))
df2$ZIP_Code <- as.factor(df2$ZIP_Code)
df2$Personal_Loan <- as.factor(df2$Personal_Loan)
```
Overview  {data-orientation=rows}
=======================================================================

## Row 1 {data-height=5}

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------
### Counts of Personal Loan, Mortgage, Securities Account, Online, and Credit Card

```{r}
df2 %>%
  group_by(Personal_Loan, Mortgage, Securities_Account, Online, CreditCard) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  gather(key = "variable", value = "value", -count) %>%
  mutate(value = ifelse(value == 1, "Accepted Loan", "Did Not Accept Loan")) %>%
  hchart(type = "column", hcaes(x = variable, y = count, group = factor(value), color = variable)) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Variable")) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  hc_title(text = "Counts of Personal Loan, Mortgage, Securities Account, Online, and Credit Card") %>% 
  hc_colors(list('#f15c80', '#7cb5ec'))
```
### Personal Loan Holders VS Securities Account and Credit Card Status  

```{r}

df2 %>%
  group_by(Personal_Loan, Securities_Account, CreditCard) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = factor(ifelse(Securities_Account == 1 & CreditCard == 1, "Have Both", ifelse(Securities_Account == 1, "Have Securities Account", ifelse(CreditCard == 1, "Have Credit Card", "Have None")))), y = count, group = factor(ifelse(Personal_Loan == 1, "Accepted Loan", "Did Not Accept Loan"), levels = c("Accepted Loan", "Did Not Accept Loan")))) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Securities Account and Credit Card Holders")) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  hc_title(text = "Counts of Personal Loan Holders based on Securities Account and Credit Card Status") 

```
### Mortgage Holders VS Securities Account and Credit Card Status  

```{r}
df2 %>%
  group_by(Mortgage, Securities_Account, CreditCard) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = factor(ifelse(Securities_Account == 1 & CreditCard == 1, "Have Both", ifelse(Securities_Account == 1, "Have Securities Account", ifelse(CreditCard == 1, "Have Credit Card", "Have None")))), y = count, group = factor(ifelse(Mortgage == 0,"Don't have Mortgage", "Have Mortgage"), levels = c("Don't have Mortgage", "Have Mortgage")))) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Securities Account and Credit Card Holders")) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  hc_title(text = "Counts of Mortgage Holders based on Securities Account and Credit Card Status") 

```
### CreditCard Holders based on Securities Account and Online Status

```{r}

df2 %>%
  group_by(CreditCard, Securities_Account, Online) %>%
  summarise(count = n()) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = factor(ifelse(Securities_Account == 1 & Online == 1, "Have Both", ifelse(Securities_Account == 1, "Have Securities Account", ifelse(CreditCard == 1, "Have Credit Card", "Have None")))), y = count, group = factor(ifelse(CreditCard == 0,"Don't have CreditCard", "Have CreditCard"), levels = c("Don't have CreditCard", "Have CreditCard")))) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Securities Account and Credit Card Holders")) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  hc_title(text = "Counts of CreditCard Holders based on Securities Account and Online Status") 
```

### {.no-padding data-height=350}

Personal Loan {data-orientation=rows}
=======================================================================

## Row 1 {data-height=5}

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------

### Personal Loan Status based on Income {data-height=600}


```{r}
df2 %>%
  group_by(Personal_Loan, Income) %>%
  summarise(count = n()) %>%
  hchart(type = "column", hcaes(x = Income, y = count, group = factor(ifelse(Personal_Loan == 0, "Did Not Accepted Loan", "Accepted Loan"), levels = c("Did Not Accepted Loan", "Accepted Loan" )))) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Income Category")) %>% 
  hc_yAxis(title = list(text = "Count")) %>% 
  hc_title(text = "Personal Loan Status based on Income")

```

### Education and Personal Loan {data-height=600}

```{r}
df2 %>%
  group_by(Education, Personal_Loan = factor(ifelse(Personal_Loan == 0, "Did Not Accepted Loan", "Accepted Loan"), levels = c( "Did Not Accepted Loan", "Accepted Loan"))) %>%
  summarise(n = n()) %>%
  group_by(Education) %>%
  mutate(pct = n/sum(n)*100) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = Education, y = pct, group = Personal_Loan), colorByPoint = TRUE) %>%
  hc_title(text = "Mortgage by Education") %>%
  hc_yAxis(title = "Percentage", labels = list(format = "{value}%")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:.1f}%</b><br>") %>%
  hc_plotOptions(column = list(stacking = "normal")) %>%
  hc_legend(layout = "horizontal", align = "center", verticalAlign = "bottom")
```

### Personal Loan by Family {data-height=600}

```{r}
df2 %>%
  group_by(Family, Personal_Loan = factor(ifelse(Personal_Loan == 0, "Did Not Accept Loan", "Accepted Loan"), levels = c("Did Not Accept Loan", "Accepted Loan"))) %>%
  summarise(n = n()) %>%
  group_by(Family) %>%
  mutate(pct = n/sum(n)*100) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = Family, y = pct, group = Personal_Loan), colorByPoint = TRUE) %>%
  hc_title(text = "Personal Loan by Education") %>%
  hc_yAxis(title = "Percentage", labels = list(format = "{value}%")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:.1f}%</b><br>") %>%
  hc_plotOptions(column = list(stacking = "normal")) %>%
  hc_legend(layout = "horizontal", align = "center", verticalAlign = "bottom")

```
### Personal Loan by Age {data-height=600}

```{r}
df2 %>%
  group_by(Personal_Loan, Age) %>%
  summarise(count = n()) %>%
  group_by(Age) %>%
  mutate(total = sum(count), 
         percentage = paste0(round(100 * count/total, 2), "%")) %>%
  hchart(type = "column", hcaes(x = Age, y = count, group = factor(ifelse(Personal_Loan == 0, "Did Not Accept Loan", "Accepted Loan"), levels = c( "Did Not Accept Loan", "Accepted Loan")))) %>% 
  hc_plotOptions(column = list(stacking = FALSE)) %>% 
  hc_legend(enabled = TRUE) %>% 
  hc_xAxis(title = list(text = "Income Category")) %>% 
  hc_yAxis(title = list(text = "Percentage")) %>% 
  hc_tooltip(pointFormat = "<b>{point.percentage}</b>") %>%
  hc_title(text = "Personal Loan Status based on Income")

```
### Distribution of ZIP Code with no Personal Loan {data-height=600}
```{r}
# subset data for Mortgage == 0
df_subset <- subset(df2, Personal_Loan == 0)

# order ZIP Code by count of Mortgage == 0
zip_order <- as.character(names(sort(table(df_subset$ZIP_Code), decreasing = TRUE)))[1:20]

# plot count of ZIP Code
ggplot(df_subset, aes(y = ZIP_Code, fill = ZIP_Code)) + 
  geom_bar() + 
  scale_fill_viridis_d(option = "mako") + 
  scale_y_discrete(limits = zip_order) + 
  theme_minimal() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(title = "Distribution of ZIP Code with no Personal Loan", style = list(fontSize='25px', 
                                                                         fontWeight= 'bold'),
       y = "ZIP Code")


```
### {.no-padding data-height=350}

Mortgage {data-orientation=rows}
=======================================================================

## Row 1 {data-height=5}

Column {.tabset .tabset-fade data-width=650}
-----------------------------------------------------------------------
### Mortgage by Income {data-height=600}

```{r}
df2 %>%
  group_by(Income, Mortgage = factor(ifelse(Mortgage == 0, "Don't have Mortgage", "Have Mortgage"), levels = c( "Don't have Mortgage", "Have Mortgage"))) %>%
  summarise(n = n()) %>%
  group_by(Income) %>%
  mutate(pct = n/sum(n)*100) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = Income, y = pct, group = Mortgage), colorByPoint = TRUE) %>%
  hc_title(text = "Mortgage by Income") %>%
  hc_yAxis(title = "Percentage", labels = list(format = "{value}%")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:.1f}%</b><br>") %>%
  hc_plotOptions(column = list(stacking = "normal")) %>%
  hc_legend(layout = "horizontal", align = "center", verticalAlign = "bottom")

```

### Education and Mortgage {data-height=600}
```{r}
df2 %>%
  group_by(Education, Mortgage = factor(ifelse(Mortgage == 0, "Don't have Mortgage", "Have Mortgage"), levels = c( "Don't have Mortgage", "Have Mortgage"))) %>%
  summarise(n = n()) %>%
  group_by(Education) %>%
  mutate(pct = n/sum(n)*100) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = Education, y = pct, group = Mortgage), colorByPoint = TRUE) %>%
  hc_title(text = "Mortgage by Education") %>%
  hc_yAxis(title = "Percentage", labels = list(format = "{value}%")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:.1f}%</b><br>") %>%
  hc_plotOptions(column = list(stacking = "normal")) %>%
  hc_legend(layout = "horizontal", align = "center", verticalAlign = "bottom")
```

### Family and Mortgage {data-height=600}

```{r}

df2 %>%
  group_by(Family, Mortgage = factor(ifelse(Mortgage == 0, "Don't have Mortgage", "Have Mortgage"), levels = c( "Don't have Mortgage", "Have Mortgage"))) %>%
  summarise(n = n()) %>%
  group_by(Family) %>%
  mutate(pct = n/sum(n)*100) %>%
  ungroup() %>%
  hchart(type = "column", hcaes(x = Family, y = pct, group = Mortgage), colorByPoint = TRUE) %>%
  hc_title(text = "Mortgage by Family") %>%
  hc_yAxis(title = "Percentage", labels = list(format = "{value}%")) %>%
  hc_tooltip(pointFormat = "<b>{point.y:.1f}%</b><br>") %>%
  hc_plotOptions(column = list(stacking = "normal")) %>%
  hc_legend(layout = "horizontal", align = "center", verticalAlign = "bottom")
```

### Distribution of ZIP Code with no Mortgage {data-height=600}
```{r}
# subset data for Mortgage == 0
df_subset <- subset(df2, Mortgage == 0)

# order ZIP Code by count of Mortgage == 0
zip_order <- as.character(names(sort(table(df_subset$ZIP_Code), decreasing = TRUE)))[1:40]

# plot count of ZIP Code
ggplot(df_subset, aes(y = ZIP_Code, fill = ZIP_Code)) + 
  geom_bar() + 
  scale_fill_viridis_d(option = "mako") + 
  scale_y_discrete(limits = zip_order) + 
  theme_minimal() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.x = element_blank(),
        axis.line.y = element_blank(),
        axis.ticks.y = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank()) +
  labs(title = "Distribution of ZIP Code with no Mortgage", style = list(fontSize='25px', 
                                                                         fontWeight= 'bold'),
       y = "ZIP Code")
```




